---
title: "hw03-MaunishUBC"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Homework 03: Using dplyr/ggplot2 to manipulate and explore data


### Installing gapminder and tidyverse
```{r  warning=FALSE}
library(gapminder)
library(tidyverse)
library(kableExtra)
```

### Task 1: Get the maximum and minimum GDP per capita for all continents
```{r}
max_min_gdpPercap <- gapminder %>% 
  group_by(continent) %>% 
  summarise(max_gdpPercap = max(gdpPercap), min_gdpPercap = min(gdpPercap))

max_min_gdpPercap %>% kable() %>% 
  kable_styling()


```
```{r}
gapminder %>% 
  ggplot(aes(continent, gdpPercap, colour = continent)) +
  scale_y_log10() +
  geom_violin() +
  geom_jitter(alpha = 0.2)
```

It is clear that Asia has the higher spread in gdpPercap

### Task 2: Look at the spread of GDP per capita (gdpPercap) within the continents

Best way to do this would be to use histograms for the GDP separately for each continent 
```{r}
gapminder %>% 
  ggplot(aes(gdpPercap)) +
  geom_histogram(bins = 50, aes(colour = continent)) + 
  scale_x_log10() + 
  facet_wrap(~ continent)
```

Also useful, would be to calculate the standard deviation of the gdpPercap for each of the continents

```{r}
gapminder %>% 
  group_by(continent) %>% 
  summarise(spread = sd(gdpPercap)) %>% 
  kable() %>% 
  kable_styling()
  
```
Asia has the highest spread in GDP per cap.

### Compute a weighted mean (weighting by population) of life expectancy for different years. 

```{r}
gapminder_weighted <- gapminder %>% 
  mutate(life_exp_wt = lifeExp/log(pop)) 

gapminder_weighted %>% 
  group_by(country) %>% 
  summarize(weighted_mean = mean(life_exp_wt)) %>% 
  kable() %>% 
  kable_styling()
```

So, what does the weighted mean spread look like?

```{r}
gapminder_weighted %>% 
  ggplot(aes(life_exp_wt)) +
  geom_histogram(bins =50, aes(fill = continent))
```
```{r}
filter(gapminder_weighted, life_exp_wt > 6) %>% 
  kable() %>% 
  kable_styling()
```

```{r}
filter(gapminder_weighted, life_exp_wt < 2) %>% 
  kable() %>% 
  kable_styling()
```

It is interesting that in case of the highest weighted mean the parameter is skewed towards country with a really small population (Iceland)

Whereas in case of the smallest values it is mostly from the 1950s and from medium and large poor countries. 


